Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'

### 2024-04-19 manuscript.tex
Version to submit

\usepackage{multirow}

@@

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

@@

San Jorge Gulf is the northernmost study area considered in this review
(Figure 1), located from Cabo dos Bahias (44°55'S) to Cabo tres Puntas
(47°06'S). It is a partially enclosed basin spanning approximately
34,000 km\^{}2 and \textasciitilde100 m of maximum depths. The gulf's
productivity supports large invertebrate and vertebrate fisheries
(Góngora et al. 2012), as well as marine mammal and seabird populations
(Yorio 2009). Two prominent frontal systems meet in the shallower
northern and southern ends of the gulf (with depths \textasciitilde40
m), which are also the places of highest productivity (Glembocki et al.
2015).

@@

Located at a more southern latitude (\textasciitilde54ºS, 68ºW), the
Beagle Channel (Figure 1) is an interoceanic passage at the southernmost
tip of South America, spanning 240 km in length and 5 km in width
(\textasciitilde1200 km\^{}2), with a depth range of 0 - 140 m. It
features complex coastlines, varying bathymetry, a prevailing
west-to-east circulation pattern, and a significant longitudinal
gradient of glacial freshwater discharge (Schloss et al. 2023). Bruno et
al. (2023) suggests that locally produced suspended particulate organic
matter (mostly composed of phytoplankton) and organic matter accumulated
in the sediments (macroalgae-originated detritus) are the primary food
sources for the marine ecosystem, as opposed to allochthonous materials.

@@

A main local stressor in the Beagle Channel is the introduction of
invasive species (Table 2). Salmonidae were introduced to Tierra del
Fuego in the 1930s. Especially, the Chinook salmon \emph{Oncorhynchus
tshawytscha} causes concern. The detection of this species in Tierra del
Fuego dates back to April 2006, and its population has been expanding
since then (Nardi et al. 2019). Being a top predator, the Chinook salmon
can compete for resources with several native predators in the Beagle
Channel (Correa and Gross 2008), but also prey on them. This is the case
for Notothenioids (\emph{Eleginops maclovinus}, \emph{Patagonotothen
tessellata}, \emph{P. cornucola}, \emph{P. sima}, \emph{Paranotothenia
magellanica}, \emph{Harpagifer bispinis}), the Atherinidae
\emph{Odontesthes smitti} and \emph{O. nigricans}, the Fuegian sprat,
and larvae of king crabs (\emph{Lithodes antarcticus} and
\emph{Paralomis granulosa}) (Fernández et al. 2010). Ciancio et al.
(2008) observed that Chinook salmon primarily feeds on sprats in the
Southern Patagonian Shelf Ecosystem area and display trophic levels
comparable to those of intermediate-sized fish and cephalopod predator
species, showing significant dietary overlap with the Magellanic penguin
\emph{Spheniscus magellanicus}. Another potential competitor for Chinook
salmon in the Beagle Channel, sharing a similar diet, is the native
Commerson's dolphin \emph{Cephalorhynchus commersonii} (Riccialdelli et
al. 2013). In the Beagle Channel, other stressors include contaminants
like metals, perfluorinated compounds, hydrocarbons, and microplastics
found in animal tissue (algae, mussel, fish, and guano) and sediments
(Duarte et al. 2011; Llorca et al. 2012). Carbon and carbohydrate levels
in Ushuaia Bay's (54°48′S, 68°18′W) surface sediments are similar to
hypertrophic ecosystems (i.e., greater input of nutrients than an
eutrophic situation), along with hydrocarbons and heavy metals linked to
port and industrial activities (Gil et al. 2011; Commendatore et al.
2012). Ferreira et al. (2021) showed that the black southern cod in
Ushuaia Bay may be exposed to endocrine-disrupting compounds from urban
and industrial pollution. Moreover, Pérez et al. (2020) and Ojeda et al.
(2021) found microplastics in the Chilean mussel \emph{Mytilus
chilensis} and the sea snail \emph{Nacella magellanica}, respectively.
These studies link pollutants to port and industrial activities of
Ushuaia; we therefore expect the area of the Beagle Channel closest to
the city to be the most affected by contaminants. Contaminants
potentially spread through the food web, magnifying the risk to higher
trophic level taxa. In the Beagle Channel and Atlantic coast of Tierra
del Fuego food webs mercury concentrations rose with benthivory, and
higher mercury was found in phytoplankton and the squat lobster. This
latter species connects pelagic and benthic habitats, so any disruption
to them could disrupt the food web (Rodriguez et al. 2022). Dodino et
al. (2022) found the highest mercury levels in Magellanic penguin
feathers from offshore colonies in Tierra del Fuego. Recently, Ushuaia's
kelp forests have also seen changes in composition along with a decrease
in biodiversity as a result of urban pollution (Kaminsky et al. 2023).
It is important to note that sea warming might be a less significant
stressor in the Beagle Channel compared to the other areas, as global
warming is melting glaciers and the influx of cold freshwater is
actually slightly cooling the seawater (Saraceno et al. 2022).

@@

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II (\textasciitilde54ºS, 59ºW), meaning
the shallow submarine plateau named Burdwood Bank with a 200 m isobath
boundary, and a deep slope that reaches 4000 m in depth, respectively
(Administración de Parques Nacionales 2022) (Figure 1). Physical
features in the plateau are fairly stable, with salinity averaging 34
all year round and temperature ranging between 4 and 8ºC (Acha et al.
2004). The plateau is surrounded by steep flanks of up to 4000 m depth,
protected by the Marine Protected Area Namuncurá - Burdwood Bank II
(32,000 km\^{}2). Intense upwelling and mixing occur in relation to the
slope, entraining deep nutrient-rich waters into the photic layer
(Matano et al. 2019), and resulting in a fairly homogeneous water column
both spatially and temporally (Matano et al. 2019).

@@

The Scotia Sea is a deep-sea basin (\textasciitilde48 - 58ºS, 50ºW),
delimited by the Drake Passage to the West and by the island complex of
the Scotia Arc to the North, East, and South, with an approximate
extension of 1.5 x 106 km\^{}2 and a depth range of 0 - 3000 m (Murphy
et al. 2006) (Figure 1). Its oceanography is dominated by the Antarctic
Circumpolar Current, which is spatially structured by frontal systems
(Whitworth 1980). The South Antarctic Circumpolar Current Front
subdivides the Scotia Sea into two biogeographic regions: the Northern
Scotia Sea, characterized by higher and more variable temperatures, and
the Southern Scotia Sea, characterized by lower and more stable
temperatures and influenced by seasonal sea ice (Raymond 2011).

@@

The Scotia Sea is a vast and heterogenous oceanic region, where
especially the zone around South Georgia island represent an area of
interest, here referred to as `Northern Scotia Sea'. The majority of
studies analyzing the stressors' effects come from this area (e.g.,
Murphy et al. (2007); Whitehouse et al. (2008); Trathan (2023)) (Table
2). The main stressor of the Scotia Sea is commercial fisheries. Krill
fishery on \emph{Euphausia superba} not only impacts the targeted
species, but also the many predators that depend on it as a food source
(Hilborn et al. 2017). Yet, data currently available from monitoring of
krill and its predators remain insufficient, hence identifying the
potential fishery impacts on the ecosystem is difficult (Trathan et al.
2021). Apart from krill fishery, two other commercial fisheries operate
in the Scotia Sea, targeting Patagonian toothfish species \emph{D.
eleginoides} and \emph{Dissostichus mawsoni}. The \emph{D. eleginoides}
stock is linked to the stock at South Georgia (`Northern Scotia Sea')
(Collins et al. 2010), while the \emph{D. mawsoni} stock is linked to
the Antarctic continental shelf (`Southern Scotia Sea') (Soeffker et al.
2022). Recently, Trathan (2023) has identified several concerns
regarding aspects of these fisheries, highlighting that it is crucial to
recognize ongoing changes such as the recovery of baleen whales and
climate change. These factors increase uncertainty for fisheries
managers, necessitating direct consideration in the management of
harvested and dependent species.

@@

In the Antarctic realm, Potter Cove is a \textasciitilde9 km2 fjord with
a depth range of 0 - 200 m located at 25 de Mayo/King George Island
(62°S, 58°W) on the South Shetland Islands of the West Antarctic
Peninsula (Figure 1). The cove, bordered by the Fourcade Glacier, is
divided into three areas: a) the internal cove, a high
glacier-influenced, soft sediment zone with a 50 m maximum depth; b) the
central cove, a mixed substrate area with low meltwater influence and an
80 m maximum depth; and c) the external cove, ice-free for 60 years with
a 185 m maximum depth and rocky substrate (Jerosch et al. 2018). Potter
Cove's high-latitude location results in variable environmental
conditions due to photoperiod length seasonality. Sea ice often covers
this area in winter (Schloss et al. 2012). With low phytoplankton
biomass, macroalgae, and microphytobenthos are likely the primary food
sources for secondary benthic production (Quartino and Boraso de Zaixso
2008).

@@

In the Weddell Sea, the main stressors are related to warming impacts
(sea warming, sea ice extent, iceberg scouring) (Table 2), where sea
warming has been substantial since the early 1980s (Turner et al. 2020).
For example, in 2017, the mean temperature for February reached 1.45°C,
the highest monthly mean ever recorded, i.e., 0.56°C above the
climatological mean (Turner et al. 2020). Warming has caused spatial and
temporal reductions in sea ice, impacting both pelagic and benthic
communities (Constable et al. 2014; Gutt et al. 2021). Since the start
of satellite imaging in 1978, sea ice extent has reached new record
lows. Some of the most pronounced impacts of decreasing sea ice are
negative changes in primary productivity (Atkinson et al. 2004) and
declines of the Antarctic krill. Moreover, reduced sea ice cover allows
increased access to krill by its predators (Kawaguchi et al. 2009),
which further contributes to its decline. Some outcomes of the krill
decline are reduced energy intake by higher trophic levels such as
humpback whales \emph{Megaptera novaeangliae} Braithwaite et al. (2015)
which rely heavily on krill for foraging, as well as reduced carbon
export due to fewer fecal pellets (Pauli et al. 2021). Many higher
trophic level species in the Weddell Sea are greatly impacted by sea ice
reductions due to loss of habitat and food source availability, such as
the Antarctic petrel \emph{Thalassoica antarctica} - one of the most
abundant seabirds in the Weddell Sea (Orgeira et al. 2021). Other
impacted sea bird species are the emperor penguin \emph{Aptenodytes
forsteri}. Reductions in sea ice constrain its population as evidenced
by recent observations in the Bellingshausen Sea region (Fretwell et al.
2023). The Arctic tern \emph{Sterna paradisaea} spends the summer in the
Weddell Sea exploiting krill swarms under receding ice edges. Thus,
continued warming is expected to gradually erode the abundance and
distribution of marine seabirds in the Weddell Sea (Orgeira et al.
2021). Many marine mammals depend on for breeding and foraging such as
the Weddell and Crabeater seals (\emph{Leptonychotes weddellii} and
\emph{Lobodon carcinophagus}) (Wege et al. 2021) and the Humpback whale.
Reduced sea ice can significantly affect the breeding success of seals
(Wege et al. 2021) and body condition of whales (Turner et al. 2020).
Notothenioid fish, with \emph{Pleuragramma antarcticum} being the most
abundant species, confront numerous stressors such as sea warming, sea
ice decline, and ocean acidification, posing significant threats to
their survival. While certain species exhibit physiological plasticity
to cope with increased oxygen demand, most notothenioid fish are
stenothermal and incapable of adjusting their metabolic functioning
(Mintenbeck et al. 2012). Projections suggest that the preferred thermal
habitat of the Antarctic toothfish, \emph{Dissostichus mawsoni}, an
important fish species in the Weddell Sea ecosystem, may contract over
the next three decades, underscoring the potential impact of global
warming on these species (Constable et al. 2014). Iceberg scouring is a
major factor in driving the high biodiversity of benthic communities in
the Weddell Sea (Gutt and Starmans 2001). Even at 600 m depths, iceberg
scouring has a strong effect on the benthic environment, disrupting the
upper layers of the seabed and removing macrofauna. This patchy
disturbance and distribution pattern occurs roughly every 200 square
meters on the Antarctic continental shelf. Global warming is predicted
to increase iceberg scouring frequency (Gutt 2001; Smale et al. 2008),
disrupting the environment (Smale and Barnes 2008) with the potential
for tipping points (Gutt et al. 2015). \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

@@

\begin{longtable}{ L{2.1cm} L{2.5cm} L{4cm} L{2.5cm} L{2cm} L{3cm} }
\hline
\textbf{Study area} & \textbf{Stressor} & \textbf{Species affected} & \textbf{Parameter affected} & \textbf{Locality} & \textbf{Reference} \\
\hline
\endhead
\endfoot
\hline
\endlastfoot

@@

In Potter Cove, a fjord-like Antarctic ecosystem, the repercussions of
climate change extend across numerous species. Noteworthy, these effects
are non-uniform within the food web, manifesting differently across its
various compartments (Cordone et al. 2020; Rodriguez et al. 2022).
Potter Cove has recently experienced frequent events of marine
heatwaves, i.e., prolonged periods of anomalously high sea surface
temperature (Oliver et al. 2018; Latorre et al. 2023). This has led to
decreases in biomasses of different planktonic functional groups (Garcia
et al. 2019; Latorre et al. 2023). Given the relatively low abundance of
phytoplankton and zooplankton's low mean degree, resulting in weak
interaction strengths between these low trophic levels and higher ones
(\textbf{Rodriguez2024?}), added to the modular configuration of the
food web (Rodriguez et al. 2022), we hypothesize that changes in these
nodes due to warming will be retained at the basal level of the pelagic
compartment and will not expand to higher trophic levels. Benthic
primary producers in Potter Cove are being influenced by the decrease in
winter sea ice cover (higher light availability), the increased levels
of sediments in the water column due to glacial melt run-off (lower
light penetration) and the newly free-ice areas available for
colonization associated to glacier retreat. The overall local effect of
climate change on macroalgae is a net increase in their production
(Deregibus et al. 2023). It has been proposed that larger diversity in
primary producers can support a more diverse food web with more
specialized consumers (Iken et al. 2023). Given the high relative
abundance and the high degree of the macroalgae functional group (Table
S1), we expect a longer hard bottom food web (i.e., number of
interactions connecting low and high-trophic levels), wider consumer
trophic niches, and a more stable benthic food web as sea ice cover
decreases and glaciers retreat due to global warming. In soft bottom
areas of the cove, multiple food web nodes are being affected by ongoing
warming effects: decrease in net primary production of benthic
microalgae (Hoffmann et al. 2019), and changes in the benthic community
biomass, distribution and composition (Sahade et al. 2015; Pasotti et
al. 2015). Given that the Potter Cove food web has low connectance and
omnivory, we suggest fragility and potential trophic cascade effects
(Marina et al. 2018), especially between species of the soft bottom
compartment (Table 3). Overall, we anticipate significant shifts in
energy fluxes, rather than alterations in the overall structure of the
Potter Cove food web, with strengthening of connections within
compartments.

@@

In the southernmost food web, the Weddell Sea, the main effect of global
warming is the decrease in sea ice extent, with reported anomalies in
the past summer seasons (Fretwell et al. 2023). Declining sea ice extent
has reduced the abundance of krill (Atkinson et al. 2004; Flores et al.
2012), and produced an increase in phytoplankton productivity (Pinkerton
et al. 2021; Isla 2023), altering the plankton community structure, and
benefiting cryptophytes over diatoms (Lin et al. 2021). Moreover,
habitat loss from sea ice decline will reduce the foraging success and
breeding sites of seabirds (e.g., snow petrel \emph{P. nivea} and
emperor penguin), decreasing their population biomassess and modifying
their distributions. The projected rise in iceberg scouring is expected
to significantly alter the biomass and community structure of
macrobenthos, which in turn will impact mid-trophic level predators such
as demersal fish (Gutt 2001; Mintenbeck et al. 2012). While we do not
anticipate large-scale topological changes in the mean food web
attributes, local extinctions could lead to such changes, particularly
affecting benthic species (Gutt and Piepenburg 2003). Given that the
impacted species---whether individually like krill, or collectively like
macrobenthos and notothenioids (e.g., \emph{P. antarcticum})---present a
mid-trophic position, high population biomasses and high degree (Table
S1), we hypothesize that significant shifts in energy fluxes will occur
(Table 3). Additionally, the food web's low proportion of omnivores
suggests reduced system resilience (Table 1), increasing the likelihood
of regime changes (Gutt et al. 2015).

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2759}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2644}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2759}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1609}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\multirow{2}{=}{\textbf{Study area}

======================= \textbf{San Jorge Gulf}} &
\multirow{2}{=}{\textbf{Stressor}

======================} &
\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4368} + 2\tabcolsep}@{}}{%
\multirow{2}{=}{\textbf{Hypothesis on food web effects} \textbar{}
=======================+ \textbar{}}} \\
 \\
& Fishery* & ↑ FW connectance

@@

↓ functional diversity & \\
& Sea warming & Shifts in FW topology

@@

↑ functional diversity & \\
\textbf{Beagle Channel} & & & \\
& Alien species & Shifts in FW topology

↑ spread of perturbations & \\
\textbf{Burdwood Bank} & & & \\
& Fishery & Shifts in main energy fluxes

↓ FW stability & \\
\textbf{Scotia Sea} & & & \\
& Sea warming & ↑ FW connectance

@@

↓ resilience & \\
\end{longtable}

>>>>

@@

@@

San Jorge Gulf is the northernmost study area considered in this review
(Figure 1), located from Cabo dos Bahias (44°55'S) to Cabo tres Puntas
(47°06'S). It is a partially enclosed basin spanning approximately
34,000 \(\text{km}^2\) and \textasciitilde100 m of maximum depths. The
gulf's productivity supports large invertebrate and vertebrate fisheries
(Góngora et al. 2012), as well as marine mammal and seabird populations
(Yorio 2009). Two prominent frontal systems meet in the shallower
northern and southern ends of the gulf (with depths \textasciitilde40
m), which are also the places of highest productivity (Glembocki et al.
2015).

@@

Located at a more southern latitude (\textasciitilde54ºS, 68ºW), the
Beagle Channel (Figure 1) is an interoceanic passage at the southernmost
tip of South America, spanning 240 km in length and 5 km in width
(\textasciitilde1200 \(\text{km}^2\)), with a depth range of 0 - 140 m.
It features complex coastlines, varying bathymetry, a prevailing
west-to-east circulation pattern, and a significant longitudinal
gradient of glacial freshwater discharge (Schloss et al. 2023). Bruno et
al. (2023) suggests that locally produced suspended particulate organic
matter (mostly composed of phytoplankton) and organic matter accumulated
in the sediments (macroalgae-originated detritus) are the primary food
sources for the marine ecosystem, as opposed to allochthonous materials.

@@

One of the main local stressors in the Beagle Channel is the
introduction of invasive species (Table 2). Salmonidae were introduced
to Tierra del Fuego in the 1930s. Especially, the Chinook salmon
\emph{Oncorhynchus tshawytscha} causes concern. The detection of this
species in Tierra del Fuego dates back to April 2006, and its population
has been expanding since then (Nardi et al. 2019). Being a top predator,
the Chinook salmon can compete for resources with several native
predators in the Beagle Channel (Correa and Gross 2008), but also prey
on them. This is the case for Notothenioids (\emph{Eleginops
maclovinus}, \emph{Patagonotothen tessellata}, \emph{P. cornucola},
\emph{P. sima}, \emph{Paranotothenia magellanica}, \emph{Harpagifer
bispinis}), the Atherinidae \emph{Odontesthes smitti} and \emph{O.
nigricans}, the Fuegian sprat, and larvae of king crabs (\emph{Lithodes
antarcticus} and \emph{Paralomis granulosa}) (Fernández et al. 2010).
Ciancio et al. (2008) observed that Chinook salmon primarily feeds on
sprats in the Southern Patagonian Shelf Ecosystem area and display
trophic levels comparable to those of intermediate-sized fish and
cephalopod predator species, showing significant dietary overlap with
the Magellanic penguin \emph{Spheniscus magellanicus}. Another potential
competitor for Chinook salmon in the Beagle Channel, sharing a similar
diet, is the native Commerson's dolphin \emph{Cephalorhynchus
commersonii} (Riccialdelli et al. 2013). In invaded ecosystems,
predation emerges as the primary driver of significant damage, often
leading to cascading effects impacting even primary producers (David et
al. 2017). In the Beagle Channel, other stressors include contaminants
like metals, perfluorinated compounds, hydrocarbons, and microplastics
found in animal tissue (algae, mussel, fish, and guano) and sediments
(Duarte et al. 2011; Llorca et al. 2012). Carbon and carbohydrate levels
in Ushuaia Bay's (54°48′S, 68°18′W) surface sediments are similar to
hypertrophic ecosystems (i.e., greater input of nutrients than an
eutrophic situation), along with hydrocarbons and heavy metals linked to
port and industrial activities (Gil et al. 2011; Commendatore et al.
2012). Ferreira et al. (2021) showed that the black southern cod in
Ushuaia Bay may be exposed to endocrine-disrupting compounds from urban
and industrial pollution. Moreover, Pérez et al. (2020) and Ojeda et al.
(2021) found microplastics in the Chilean mussel \emph{Mytilus
chilensis} and the sea snail \emph{Nacella magellanica}, respectively.
These studies link pollutants to port and industrial activities of
Ushuaia; we therefore expect the area of the Beagle Channel closest to
the city to be the most affected by contaminants. Contaminants
potentially spread through the food web, from lower to higher trophic
level taxa. In the Beagle Channel and Atlantic coast of Tierra del Fuego
food webs mercury concentrations rose with benthivory, and higher
mercury was found in phytoplankton and the squat lobster. This latter
species connects pelagic and benthic habitats, so any disruption to them
could disrupt the food web (Rodriguez et al. 2022). Dodino et al. (2022)
found the highest mercury levels in Magellanic penguin feathers from
offshore colonies in Tierra del Fuego. Recently, Ushuaia's kelp forests
have also seen changes in composition along with a decrease in
biodiversity as a result of urban pollution (Kaminsky et al. 2023). It
is important to note that sea warming might be a less significant
stressor in the Beagle Channel compared to the other areas, as global
warming is melting glaciers and the influx of cold freshwater is
actually slightly cooling the seawater (Saraceno et al. 2022).

@@

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II (\textasciitilde54ºS, 59ºW), meaning
the shallow submarine plateau named Burdwood Bank with a 200 m isobath
boundary, and a deep slope that reaches 4000 m in depth, respectively
(Administración de Parques Nacionales 2022) (Figure 1). Physical
features in the plateau are fairly stable, with salinity averaging 34
all year round and temperature ranging between 4 and 8ºC (Acha et al.
2004). The plateau is surrounded by steep flanks of up to 4000 m depth,
protected by the Marine Protected Area Namuncurá - Burdwood Bank II
(32,000 \(\text{km}^2\)). Intense upwelling and mixing occur in relation
to the slope, entraining deep nutrient-rich waters into the photic layer
(Matano et al. 2019), and resulting in a fairly homogeneous water column
both spatially and temporally (Matano et al. 2019).

@@

The Scotia Sea is a deep-sea basin (\textasciitilde48 - 58ºS, 50ºW),
delimited by the Drake Passage to the West and by the island complex of
the Scotia Arc to the North, East, and South, with an approximate
extension of 1.5 x 106 \(\text{km}^2\) and a depth range of 0 - 3000 m
(Murphy et al. 2006) (Figure 1). Its oceanography is dominated by the
Antarctic Circumpolar Current, which is spatially structured by frontal
systems (Whitworth 1980). The South Antarctic Circumpolar Current Front
subdivides the Scotia Sea into two biogeographic regions: the Northern
Scotia Sea, characterized by higher and more variable temperatures, and
the Southern Scotia Sea, characterized by lower and more stable
temperatures and influenced by seasonal sea ice (Raymond 2011).

@@

The Scotia Sea is a vast and heterogenous oceanic region, where
especially the zone around South Georgia island represent an area of
interest, here referred to as `Northern Scotia Sea'. The majority of
studies analyzing the stressors' effects come from this area (e.g.,
Murphy et al. (2007); Whitehouse et al. (2008); Trathan (2023)) (Table
2). The main stressor of the Scotia Sea is commercial fisheries. Krill
fishery on \emph{Euphausia superba} not only impacts the targeted
species, but also the many predators that depend on it as a food source
(Hilborn et al. 2017). Yet, data currently available from monitoring of
krill and its predators remain insufficient, hence identifying the
potential fishery impacts on the ecosystem is difficult (Trathan et al.
2021). Apart from krill fishery, two other commercial fisheries operate
in the Scotia Sea, targeting Patagonian toothfish species \emph{D.
eleginoides} and \emph{Dissostichus mawsoni}. The \emph{D. eleginoides}
stock is linked to the stock at South Georgia (`Northern Scotia Sea')
(Collins et al. 2010), while the \emph{D. mawsoni} stock is linked to
the Antarctic continental shelf (`Southern Scotia Sea') (Soeffker et al.
2022). Recently, Trathan (2023) has identified several concerns
regarding aspects of these fisheries, highlighting that it is crucial to
recognize ongoing changes such as the recovery of baleen whales and
climate change. These factors increase the uncertainty for fishery
managers, necessitating direct consideration in the management of
harvested and dependent species.

@@

In the Antarctic realm, Potter Cove is a \textasciitilde9
\(\text{km}^2\) fjord with a depth range of 0 - 200 m located at 25 de
Mayo/King George Island (62°S, 58°W) on the South Shetland Islands of
the West Antarctic Peninsula (Figure 1). The cove, bordered by the
Fourcade Glacier, is divided into three areas: a) the internal cove, a
high glacier-influenced, soft sediment zone with a 50 m maximum depth;
b) the central cove, a mixed substrate area with low meltwater influence
and an 80 m maximum depth; and c) the external cove, ice-free for 60
years with a 185 m maximum depth and rocky substrate (Jerosch et al.
2018). Potter Cove's high-latitude location results in variable
environmental conditions due to photoperiod length seasonality. Sea ice
often covers this area in winter (Schloss et al. 2012). With low
phytoplankton biomass, macroalgae, and microphytobenthos are likely the
primary food sources for secondary benthic production (Quartino and
Boraso de Zaixso 2008).

@@

In the Weddell Sea, the main stressors are related to warming impacts
(sea warming, sea ice extent, iceberg scouring) (Table 2), where sea
warming has been substantial since the early 1980s (Turner et al. 2020).
For example, in 2017, the mean temperature for February reached 1.45°C,
the highest monthly mean ever recorded, i.e., 0.56°C above the
climatological mean (Turner et al. 2020). Warming has caused spatial and
temporal reductions in sea ice, impacting both pelagic and benthic
communities (Constable et al. 2014; Gutt et al. 2021). Since the start
of satellite imaging in 1978, sea ice extent has reached new record
lows. Some of the most pronounced impacts of decreasing sea ice are
negative changes in primary productivity (Atkinson et al. 2004) and
declines of the Antarctic krill. Moreover, reduced sea ice cover allows
increased access to krill by its predators (Kawaguchi et al. 2009),
which further contributes to its decline. Some outcomes of the krill
decline are reduced energy intake by higher trophic levels such as
humpback whales \emph{Megaptera novaeangliae} Braithwaite et al. (2015)
which rely heavily on krill for foraging, as well as reduced carbon
export due to fewer fecal pellets (Pauli et al. 2021). Many higher
trophic level species in the Weddell Sea are greatly impacted by sea ice
reductions due to loss of habitat and food source availability, such as
the Antarctic petrel \emph{Thalassoica antarctica} - one of the most
abundant seabirds in the Weddell Sea (Orgeira et al. 2021). Other
impacted sea bird species are the emperor penguin \emph{Aptenodytes
forsteri}. Reductions in sea ice constrain its population as evidenced
by recent observations in the Bellingshausen Sea region (Fretwell et al.
2023). The Arctic tern \emph{Sterna paradisaea} spends the summer in the
Weddell Sea exploiting krill swarms under receding ice edges. Thus,
continued warming is expected to gradually erode the abundance and
distribution of marine seabirds in the Weddell Sea (Orgeira et al.
2021). Many marine mammals depend on for breeding and foraging such as
the Weddell and Crabeater seals (\emph{Leptonychotes weddellii} and
\emph{Lobodon carcinophagus}) (Wege et al. 2021) and the Humpback whale.
Reduced sea ice can significantly affect the breeding success of seals
(Wege et al. 2021) and body condition of whales (Turner et al. 2020).
Notothenioid fish, with \emph{Pleuragramma antarcticum} being the most
abundant species, confront numerous stressors such as sea warming, sea
ice decline, and ocean acidification, posing significant threats to
their survival. While certain species exhibit physiological plasticity
to cope with increased oxygen demand, most notothenioid fish are
stenothermal and incapable of adjusting their metabolic functioning
(Mintenbeck et al. 2012). Projections suggest that the preferred thermal
habitat of the Antarctic toothfish, \emph{Dissostichus mawsoni}, an
important fish species in the Weddell Sea ecosystem, may contract over
the next three decades, underscoring the potential impact of global
warming on these species (Constable et al. 2014). Iceberg scouring is a
major factor in driving the high biodiversity of benthic communities in
the Weddell Sea (Gutt and Starmans 2001). Even at 600 m depths, iceberg
scouring has a strong effect on the benthic environment, disrupting the
upper layers of the seabed and removing macrofauna. This patchy
disturbance and distribution pattern occurs roughly every 200 square
meters on the Antarctic continental shelf. Global warming is predicted
to increase iceberg scouring frequency (Gutt 2001; Smale et al. 2008),
disrupting the environment (Smale and Barnes 2008) with the potential
for tipping points (Gutt et al. 2015).

@@

\begin{longtable}{ L{2.1cm} L{2.5cm} L{4cm} L{2.5cm} L{2cm} L{3cm} }
\hline
\textbf{Study area} & \textbf{Stressor} & \textbf{Species affected} & \textbf{Parameter affected} & \textbf{Locality} & \textbf{Reference} \\
\hline
\endhead
\endfoot
\hline
\endlastfoot

@@

In Potter Cove, a fjord-like Antarctic ecosystem, the repercussions of
climate change extend across numerous species. Noteworthy, these effects
are non-uniform within the food web, manifesting differently across its
pelagic and benthic compartments. Potter Cove has recently experienced
frequent events of marine heatwaves, i.e., prolonged periods of
anomalously high sea surface temperature (Oliver et al. 2018; Latorre et
al. 2023). This has led to decreases in biomasses of different
planktonic functional groups (Garcia et al. 2019; Latorre et al. 2023).
Given the relatively low abundance of phytoplankton and zooplankton low
mean degree (Table S1), resulting in weak interaction strengths between
these low trophic levels and higher ones (Rodriguez and Saravia 2024),
added to the modular configuration of the food web (Rodriguez et al.
2022), we hypothesize that changes in these nodes due to warming will be
retained at the lower trophic levels of the pelagic compartment and will
not expand to higher trophic levels. Macroalgae, important benthic
primary producers, are being influenced by the decrease in winter sea
ice cover (higher light availability), the increased levels of sediments
in the water column due to glacial melt run-off (lower light
penetration) and the newly free-ice areas available for colonization
associated to glacier retreat. The local effect of climate change on
macroalgae is a net increase in their production and macroalgal detritus
(Braeckman et al. 2019; Deregibus et al. 2023; Iken et al. 2023). On the
other hand, increased glacial runoff lead to a decrease in net primary
production of benthic microalgae (Hoffmann et al. 2019); reduced
secondary benthic production; changes on the benthic community
composition and an expansion of its distribution towards newly free-ice
areas, specially filter and deposit feeders species (Sahade et al. 2015;
Pasotti et al. 2015; Braeckman et al. 2024). In an increasing glacial
melt disturbance scenario, it's expected a shift in the food sources of
benthic species towards grazing on macroalgal and their detritus
(Braeckman et al. 2024). It has been proposed that larger diversity in
primary sources can support a more diverse food web with more
specialized consumers (Iken et al. 2023). Given the high relative
abundance and the high degree of the macroalgae functional group (Table
S1), we expect a longer benthic food web (i.e., number of interactions
connecting low and high-trophic levels), wider consumer trophic niches,
meaning an increase in omnivory, and a more stable benthic food web as
sea ice cover decreases and glaciers retreat due to global warming. As a
net effect of climate change on the overall Potter Cove food web, we
expect a significant shift in energy fluxes, rather than drastic
alterations in the topological structure, affecting the transfer of
energy from lower to higher trophic levels.

@@

In the southernmost food web, the Weddell Sea, the main effect of global
warming is the decrease in sea ice extent, with reported anomalies in
the past summer seasons (Fretwell et al. 2023). Declining sea ice extent
has reduced the abundance of krill (Atkinson et al. 2004; Flores et al.
2012), and produced an increase in phytoplankton productivity (Pinkerton
et al. 2021; Isla 2023), altering the plankton community structure, and
benefiting cryptophytes over diatoms (Lin et al. 2021). Moreover,
habitat loss from sea ice decline will reduce the foraging success and
breeding sites of seabirds (e.g., snow petrel \emph{P. nivea} and
emperor penguin), decreasing their population biomassess and modifying
their distributions. The projected rise in iceberg scouring is expected
to significantly alter the biomass and community structure of
macrobenthos, which in turn will impact mid-trophic level predators such
as demersal fish (Gutt 2001; Mintenbeck et al. 2012). While we do not
anticipate large-scale topological changes affecting mean food web
attributes, local extinctions may produce changes in the interactions,
especially impacting benthic species (Gutt and Piepenburg 2003). Given
that the impacted species---whether individually like krill, or
collectively like macrobenthos and notothenioids (e.g., \emph{P.
antarcticum})---present a mid-trophic position, high population
biomasses and high degree (Table S1), we hypothesize that significant
shifts in energy fluxes will occur (Table 3). Additionally, the food
web's low proportion of omnivores suggests reduced system resilience
(Table 1), increasing the likelihood of regime changes (Gutt et al.
2015).

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3194}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Study area}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Stressor}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Hypothesis on food web effects}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{San Jorge Gulf} & & \\
& Fishery* & ↑ FW connectance

@@

↓ functional diversity \\
& Sea warming & Shifts in FW topology

@@

↑ functional diversity \\
\textbf{Beagle Channel} & & \\
& Alien species & Shifts in FW topology

↑ spread of perturbations \\
\textbf{Burdwood Bank} & & \\
& Fishery & Shifts in main energy fluxes

↓ FW stability \\
\textbf{Scotia Sea} & & \\
& Sea warming & ↑ FW connectance

@@

↓ resilience \\
\end{longtable}
